\(Q\)-Boson model and relations with integrable hierarchies

### 2024-06-24 draft.tex
finish new version

Define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\) and
\(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) by
\begin{equation}
  t_q = \frac{1}{q}\sum_{j=1}^N x_j^p\qquad 
  t'_{-q} = \frac{1}{q}\sum_{j=1}^N y_j^p\; , 
\end{equation}
where \(p_q(\bm{x}) = q t_q\) are power sums. One can write the inner
product in terms of these coordinates, that is
\begin{equation}
  \mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
This expression is known to be a tau function for \(M, N\to \infty\).
As such, utilizing the free fermions representation~\cite{Alexandrov:2012tr}, it can be written as
\begin{equation}
  \lim_{M,N\to \infty}\mathcal{I}(N, M|\bm{t}, \bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}
It is a tau function of the Toda hierarchy with trivial element
\(\mathbb{1} \in GL(\infty)\), and it is nothing but the Cauchy's
identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_{-m} \right) \; .
\end{equation}

@@

\begin{remark}
This expression also reveals something interesting. We can formulate
this problem in terms of partition statess defined in the phase
model. Alternatively, we can utilize the conventional vertex operator
construction rather than the q-deformed version proposed by
Jing~\cite{Jing1991, Jing1995}, which is much more challenging to
handle.
\end{remark}

@@

We also wrote these formulas in terms of simpler orthogonal
polynomials. Firstly, we showed that these scalar products cannot be
naively written as Toda tau functions. However, using some known
expansions, we expanded the inner products in different forms. In one
of these expansions, we defined a new set of coordinates, making the
scalar product of q-bosons Toda tau functions with respect to these
new coordinates. This is the main result of this section.

>>>>

Define two sets of Miwa coordinates \(\bm{t} = (t_1, t_2, \dots)\) and
\(\bm{t}' = (t'_{-1}, t'_{-2}, \dots)\) as
\begin{equation}
  t_q = \frac{1}{q}\sum_{j=1}^N x_j^p\qquad 
  t'_{-q} = \frac{1}{q}\sum_{j=1}^N y_j^p\; , 
\end{equation}
where \(p_q(\bm{x}) = q t_q\) are power sums. One can write the inner
product in terms of these coordinates, that is
\begin{equation}
  \mathcal{I}(N, M|\bm{t}, \bm{t}') = \sum_{\lambda\subseteq [N,M]} s_\lambda(\bm{t}) s_\lambda(\bm{t}') \; .
\end{equation}
This expression is known to be a tau function for \(M, N\to \infty\).
As such, utilizing the free fermions representation~\cite{Alexandrov:2012tr}, it can be written as
\begin{equation}
  \lim_{M,N\to \infty}\mathcal{I}(N, M|\bm{t}, \bm{t}')
 = \bra{\bm{0}} e^{\bm{J}_+(\bm{t})} e^{-\bm{J}_-(\bm{t}')} \ket{\bm{0}}\; .
\end{equation}
It is a tau function of the Toda hierarchy with trivial element
\(\mathbb{1} \in GL(\infty)\), and it is nothing but the Cauchy's
identity
\begin{equation}
  \sum_{\lambda } s_{\lambda}(\bm{t}) s_{\lambda}(\bm{t}')
    = \exp \left( \sum_{m\geq1} m t_m t'_{-m} \right) \; .
\end{equation}

@@

\begin{remark}
This expression also reveals something interesting. We can formulate
this problem in terms of partition states defined in the phase
model. Alternatively, we can utilize the conventional vertex operator
construction rather than the q-deformed version proposed by
Jing~\cite{Jing1991, Jing1995}, which is much more challenging to
handle.
\end{remark}

@@

We also write these formulas in terms of simpler orthogonal
polynomials. Firstly, we show that these scalar products cannot be
naively written as Toda tau functions. However, using some known
expansions, we expand the inner products in different forms. In one
of these expansions, we define a new set of coordinates, making the
scalar product of q-bosons Toda tau functions with respect to these
new coordinates. This is the main result of this section.
